Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2022-12-06 SuppMat.tex
First resubmission

\maketitle
\section{Minkowski's structure metric $q_2$ in an extended parameter space}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{lam_256-128-64-51.pdf}
    \includegraphics[width=0.4\textwidth]{gamma15_deltaTheta.pdf}
    \caption{LEFT PANEL. Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and 
    $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text). 
    The time interval during which $q_2 \approx 1$ signals the emergence of rivulets, also for $\lambda = 64 h_0$ which is comparable with the spinodal wavelength $\lambda_s \approx 70 h_0$. RIGHT PANEL. Comparison of the Minkowski's structure metric $q_2$ for $\delta\theta=5^{\circ}$ and $\delta\theta=10^{\circ}$, 
    $\Gamma = 15$ and $\lambda = 256 h_0$.}
    \label{fig:q2_difflambda}
\end{figure}
%\begin{figure}
%    \centering
%    \includegraphics[width=0.7\textwidth]{different_contrasts.pdf}
%    \caption{Comparison of the Minkowski's structure metric $q_2$ for different values of %$\delta\theta$, 
%    $\Gamma = 1.5$ and $\lambda = 256 h_0$.}
%    \label{fig:dif_contrast}
%\end{figure}
\noindent In this section we test the robustness of the observation of the rivulet 
state over a wider parameter space and in particular at changing: 1) the 
pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
To this aim we have run simulations with $\Gamma=1.5$, $\lambda/h_0 = 64, 128$, and with $\Gamma = 15$, $\lambda = 256 h_0$, $\delta \theta = 5^{\circ}$, respectively.
In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose 
increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for 
the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported 
for comparison).
Analogously, in the right panel, we report $q_2$ as a function of time for  
$\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over 
a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e. the rivulets lifetime decreases with 
$\delta \theta$. This was somehow expected, since obviously (and also in the static case) 
the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{KonnurPRL2000}).

@@

The droplet shape is determined by minimization of the total interfacial energy:

@@

Eq.~(\ref{eq:energy}) can be rewritten as:

@@

$A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives:

@@

\frac{\pi h}{6}(3a^2 + h^2) \approx \frac{\pi}{2} a^2 h =  \frac{h_0 \lambda^2}{2},    

@@

energy (that we indicate now as $E(a)$ to lighten the notation) reads:

@@

energy is first order) and we finally get:

@@

  a_0 = \left[\frac{2h_0\lambda^4}{\pi^2(1-\cos \theta_m)}\right]^{1/6};

@@

to this order the droplet contact angle $\tan (\theta_d/2) = h/a$ reads

@@

(\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$;
at the first order in $\delta \theta$ we get:

@@

such that the correction to the droplet contact angle provides:

>>>>

\maketitle
\section{Minkowski's structure metric $q_2$ in an extended parameter space}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{SupMatFig_1.pdf}
    \includegraphics[width=0.4\textwidth]{SupMatFig_2.pdf}
    \caption{LEFT PANEL. Minkowski's structure metric $q_2$ for three different wavelengths $\lambda=64 h_0$, $\lambda=128 h_0$ and 
    $\lambda=256 h_0$ (as in figure 4 of the main text) and for $\Gamma=1.5$ (all other parameters are as in the main text). 
    The time interval during which $q_2 \approx 1$ signals the emergence of rivulets, also for $\lambda = 64 h_0$ which is comparable with the spinodal wavelength $\lambda_s \approx 70 h_0$. RIGHT PANEL. Comparison of the Minkowski's structure metric $q_2$ for $\delta\theta=5^{\circ}$ and $\delta\theta=10^{\circ}$, 
    $\Gamma = 15$ and $\lambda = 256 h_0$.}
    \label{fig:q2_difflambda}
\end{figure}
%\begin{figure}
%    \centering
%    \includegraphics[width=0.7\textwidth]{different_contrasts.pdf}
%    \caption{Comparison of the Minkowski's structure metric $q_2$ for different values of %$\delta\theta$, 
%    $\Gamma = 1.5$ and $\lambda = 256 h_0$.}
%    \label{fig:dif_contrast}
%\end{figure}
\noindent In this section we test the robustness of the observation of the rivulet 
state over a wider parameter space and in particular at changing: 1) the 
pattern wavelength $\lambda$ and 2) the heterogeneity amplitude $\delta \theta$.
To this aim we have run simulations with $\Gamma=1.5$, $\lambda/h_0 = 64, 128$, and with $\Gamma = 15$, $\lambda = 256 h_0$, $\delta \theta = 5^{\circ}$, respectively.
In Fig.~\ref{fig:q2_difflambda} (left panel) we report the measurements of the Minkowski's structure metric $q_2$, whose 
increase from zero up to a plateauing value of $q_2 \approx 1$ indicates the emergence of rivulets, for 
the different $\lambda$'s ($\lambda=256 h_0$, which is the value considered in the main text, is also reported 
for comparison).
Analogously, in the right panel, we report $q_2$ as a function of time for  
$\delta \theta = 5^{\circ}$  and $\delta \theta = 10^{\circ}$ (the value used in the main text). 
We see that the $q_2$ metric attains the value $q_2 \approx 1$, signalling the rivulet state, albeit over 
a time interval shorter than for $\delta \theta = 10^{\circ}$, i.e. the rivulets lifetime decreases with 
$\delta \theta$. This was somehow expected, since obviously (and also in the static case) 
the patterning looses effectiveness as the contact angle mismatch is reduced (see, e.g.~\cite{KonnurPRL2000}).

@@

The droplet shape is determined by minimization of the total interfacial energy

@@

Eq.~(\ref{eq:energy}) can be rewritten as

@@

$A$ and (\ref{eq:contact2}), neglecting higher than second order terms, in (\ref{eq:energyoung}) gives

@@

\frac{\pi h}{6}(3a^2 + h^2) \approx \frac{\pi}{2} a^2 h =  \frac{h_0 \lambda^2}{2}    

@@

energy (that we indicate now as $E(a)$ to lighten the notation) reads

@@

energy is first order) and we finally get

@@

  a_0 = \left[\frac{2h_0\lambda^4}{\pi^2(1-\cos \theta_m)}\right]^{1/6}.

@@

To this order the droplet contact angle $\tan (\theta_d/2) = h/a$ reads

@@

(\ref{eq:minim}) in the form $a = a^{(0)} + \delta \theta a^{(1)} + \dots$, where $a^{(0)} \equiv a_0$.
At the first order in $\delta \theta$ we get

@@

such that the correction to the droplet contact angle provides

### 2022-12-06 letter_to_editor.tex
First resubmission

\noindent Dear Editor,\\
\\
\noindent We gratefully thank you for sending us the reports on our manuscript. We are glad to see 
that the Referees agree that the work is interesting and scientifically relevant, with one of them 
(Referee B) inclined to accept it. 
We have made a major revision taking into account all Referees' reasonable remarks, including novel 
results from new simulations and theoretical derivations (see the new Supplementary Material).
Let us, please, stress further how and why we deem that the present paper meets the Physical
Review Letters criteria of impact and innovation (interest, as we already noted, was acknowledged by the Referees).
We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
suggested how to exploit it to control the dewetting morphologies. 
We think that the impact of this discovery is out of question, as witnessed by the various publications, also on in PRL on dewetting and closely related problems
(just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
``Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes'', {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); 
C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
What it is questionable here, instead, is the reliability of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.
On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
doubts and concerns. 
Therefore, we would like to bring to your attention again our manuscript in its new form, confident that it might now be suitable to appear in Physical Review Letters.\\
\\
With our best regards,\\
the Authors.

>>>>

\noindent Dear Editor,\\
\\
\noindent We gratefully thank you for sending us the reports on our manuscript. We are glad to see 
that the Referees agree that the work is interesting and scientifically relevant, with one of them 
(Referee B) inclined to accept it. 
\textcolor{red}{In particular, we would like to underline that not a single 
concern was raised on the physics, namely on the reliability of the results.
On the other hand, we acknowledge that certain points were not explained and/or discussed adequately, so we have put a lot of effort in clarifying all
doubts and concerns.}
We have made, therefore, a major revision taking into account all Referees' reasonable remarks, including novel 
results from new simulations and theoretical derivations (see the new Supplementary Material).
Let us, please, stress further how and why we deem that the present paper meets the Physical
Review Letters criteria of impact and innovation (interest, as we already noted, was acknowledged by the Referees).
We have found a novel effect in the dewetting on substrates with properly tailored time dependent properties, and 
suggested how to exploit it to control the dewetting morphologies, 
\textcolor{red}{even proposing, motivated by one Referee's suggestion, a possible experimental realisation.}
We think that the impact of this discovery is out of question, as witnessed also by the various publications, also on in PRL on dewetting and closely related problems
(just to cite but a few recent ones: J.C. Fern\'andez-Toledano et al, ``How wettability controls nanoprinting'',
{\it Phys. Rev. Lett.} {\bf 124} 224503 (2020); A.A. Pahlavan et al,
``Evaporation of binary-mixture liquid droplets: the formation of picoliter pancakelike shapes'', {\it Phys. Rev. Lett.} {\bf 127}, 024501 (2021); 
C. Clavaud et al, ``Modification of the fluctuation dynamics of ultrathin wetting films'', {\it Phys. Rev. Lett.} {\bf 126}, 228004 (2021)).
What it is somehow questionable, instead, is the trustworthiness of some of the Referees in judging a paper on microscale multiphase fluid dynamics: one of 
them (Referee A) ignores the concept of disjoining pressure, confuses dewetting with prewetting and concepts from the physics of free surface flows to wetting.
The main ``argument'' of Referee C against publication is that he/she does not see ``how would the transport of a fluid by rivulets 
surpass the transport by droplets''; but the paper is about controlling morphologies, not about transport.\\
\textcolor{red}{We believe, then, that our manuscript, in its new version, deserves to be brought again to your attention as a contribution to Physical Review Letters, 
confident that it might be now suitable of publication.}\\
\\
With our best regards,\\
the Authors.

### 2022-12-06 main_AS_final.tex
First resubmission

by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801}.

@@

where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave~\footnote{\textcolor{red}{A laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
The external stimulus, our moving pattern, can be realized by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}}. 

@@

    \includegraphics[width=0.45\textwidth]{Figures/No_vel_render2.png}
    \caption{Stationary film thickness field ($t>4t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 

@@

    \includegraphics[width=0.45\textwidth]{Rupture_times_Gamma_theta_20.pdf}
    %\includegraphics[width=0.45\textwidth]{Figures/Rupture_times_Gamma.pdf}
    % \includegraphics[width=0.45\textwidth]{Figures/rupture_times_all_lambda_v0v10.pdf}

@@

    \includegraphics[width=0.45\textwidth]{Static_new_theta20t0.pdf}

@@

    \includegraphics[width=0.45\textwidth]{lam_2_inset_theta20.pdf}

@@

In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge. 
The outer sum represents an ensemble average over the $N$ points in the set.}. The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction. 
Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
\begin{figure}
    \centering
    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
    \includegraphics[width=0.45\textwidth]{lig_lifetime_Gamma_theta20.pdf}
    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with 
    the theoretical prediction, Eq.~(\ref{eq:rivlt}).
    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup. 
These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.

We argue that the emergence of rivulets is controlled by the competition of two characteristic velocities: the pattern wave speed, $v_{\theta}$, and the retraction speed, $U_{\Theta}$, introduced in Eq.~(\ref{eq:taur_l1}). 
If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the transport due to the contact angle field and thus droplets from. However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state. \textcolor{red}{It appears, therefore, natural to consider the ratio of these two velocities, 
$\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.}
%Recalling the expression (\ref{eq:t0}) for the reference velocity $v_0$, %we define the parameter 
%\begin{equation}\label{eq:vel_ratio}
%    \Gamma = \frac{v_{\theta}}{U_{\Theta}} = \frac{3\lambda h_0^3 %q_0^4}{\Theta^3}\chi 
%\end{equation}
%as the ratio of these two velocities,
%where $\chi \equiv v_{\theta}/v_0$. 
We see from Fig.~\ref{fig:msm_q2} that indeed rivulets form only for $\Gamma > 1$. 
Moreover, the larger $\Gamma$, the more stable the rivulets are; in other words, the rivulet life-time, $\tau_{\text{riv}}$, that can be conventionally taken as the time at which the drop of $q_2$ occurs, grows with $v_{\theta}$ (see Fig.~\ref{fig:stab_ligs_lam2}).
The rivulet itself is, in fact, prone to dewetting, with the liquid accumulating over patches around contact angle minima. However, as the pattern moves, the instability is tamed due to configurations whereby higher contact angle regions underlie height field maxima, thus tending to revert the fluid flow.
Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, this should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration. Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore 
$\Delta h(t)/\Delta h_0 \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$. 
The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields
\begin{equation}\label{eq:rivlt}
    \tau_{\text{riv}} \sim \alpha \log(v_{\theta}) \propto t_0 \log(\Gamma). 
\end{equation}
This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.

\noindent {\it Conclusions.}
We have presented results, from numerical simulations and theoretical analysis, on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
Studying how the stability of the film depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit. 
In the time-dependent case, the rupture times are in general longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behaviour is observed as the wavelength decreases. 
A theoretical explanation has been provided for all these various regimes.
Furthermore, we showed that, at increasing the wettability wave speed, a transition occurs in the dewetting morphology from a multi-droplet to a metastable multi-rivulet state. 
A dimensionless parameter, $\Gamma$, controlling the transition has been identified in the ratio of the pattern speed and the typical film retraction speed over the substrate. 
The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that has been captured by means of phenomenological arguments.
On a broader perspective, our work suggests that switchable substrates offer a new avenue to control thin film dewetting, with obviously relevant implications, for instance, for open microfluidic devices, and paves the way to future studies in this direction, exploiting more complex and dedicated space-time dependencies.

\noindent {\it Acknowledgements}. SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11.

>>>>

by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.

@@

where $q_{\theta} = 2\pi/\lambda$, i.e.~the pattern evolves in time as a plane wave.
\textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 
The external stimulus, our moving pattern, can be realized, e.g., by using a digital multimirror device (DMD)~\cite{doi:10.1021/jp301092y,doi:10.1126/science.aax8760}.}

@@

    \includegraphics[width=0.45\textwidth]{Figure_1.png}
    \caption{Stationary film thickness field ($t>t_0$) showing the formation of droplets. The color map indicates the contact angle pattern 

@@

    \includegraphics[width=0.45\textwidth]{Figure_2.pdf}

@@

    \includegraphics[width=0.45\textwidth]{Figure_3.pdf}

@@

    \includegraphics[width=0.45\textwidth]{Figure_4.pdf}

@@

In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge. 
The outer sum represents an ensemble average over the $N$ points in the set.}. The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction. 
Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
\begin{figure}
    \centering
    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
    \includegraphics[width=0.45\textwidth]{Figure_5.pdf}
    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with 
    the theoretical prediction, Eq.~(\ref{eq:rivlt}).
    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup. 
These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.

We argue that the emergence of rivulets is controlled by the competition of two characteristic velocities: the pattern wave speed, $v_{\theta}$, and the retraction speed, $U_{\Theta}$, introduced in Eq.~(\ref{eq:taur_l1}). 
If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the transport due to the contact angle field and thus droplets from. However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state. \textcolor{red}{It appears, therefore, natural to consider the ratio of these two velocities, 
$\Gamma \equiv v_{\theta}/U_{\theta}$, as the discriminating parameter.}
%Recalling the expression (\ref{eq:t0}) for the reference velocity $v_0$, %we define the parameter 
%\begin{equation}\label{eq:vel_ratio}
%    \Gamma = \frac{v_{\theta}}{U_{\Theta}} = \frac{3\lambda h_0^3 %q_0^4}{\Theta^3}\chi 
%as the ratio of these two velocities,
%where $\chi \equiv v_{\theta}/v_0$. 
We see from Fig.~\ref{fig:msm_q2} that indeed rivulets form only for $\Gamma > 1$. 
Moreover, the larger $\Gamma$, the more stable the rivulets are; in other words, the rivulet life-time, $\tau_{\text{riv}}$, that can be conventionally taken as the time at which the drop of $q_2$ occurs, grows with $v_{\theta}$ (see Fig.~\ref{fig:stab_ligs_lam2}).
The rivulet itself is, in fact, prone to dewetting, with the liquid accumulating over patches around contact angle minima. However, as the pattern moves, the instability is tamed due to configurations whereby higher contact angle regions underlie height field maxima, thus tending to revert the fluid flow.
Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, this should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration. Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore 
$\Delta h(t)/\Delta h_0 \sim \alpha (\lambda/v_{\theta})e^{\alpha t}$. 
The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields
\begin{equation}\label{eq:rivlt}
    \tau_{\text{riv}} \sim \alpha \log(v_{\theta}) \propto t_0 \log(\Gamma). 
This logarithmic dependence is indeed observed in the numerical data as shown in Fig.~\ref{fig:stab_ligs_lam2}.\\
\textcolor{red}{Before concluding, we would like to envisage a possible realization of a 
dewetting experiment on a switchable substrate of the type modelled by the spatio-temporal 
contact angle (\ref{eq:sinetheta}). One may think of a thin liquid film cast on a light responsive substrate~\cite{IchimuraEtAl_Science2000}, under the action of controlled 
external stimuli (a light emitter). 
An ideal candidate could be a digital multimirror device (DMD).
This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)
with spatial resolution of $\approx 10 \times 10 \, \mu \text{m}^2$ (the size of a pixel).
Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
of polystyrene deposited on an oxidized silicon wafer,  
we evaluate the retraction speed 
$U_{\Theta} = \frac{\Theta^3 \gamma}{9 \mu}$ to be 
$U_{\Theta} \approx 10^{-2} \, \mu \text{m}/\text{s}$. 
The (minimum) {\it pattern speed} can be 
estimated, from the pixel size with frame rate $\sim 1 s^{-1}$, as 
$v_{\theta} \sim 10 \, \mu \text{m}/\text{s}$, that would give $\Gamma \sim 10^3$, i.e. well within the rivulet regime ($\Gamma > 1$). We note, also, that both $U_{\Theta}$ and $v_{\theta}$ can be widely modulated: the former, playing either with the (temperature and molecular weight dependent) viscosity 
or tailoring the substrate to making it more hydrophobic (i.e. increasing the contact angle); the latter, tuning
the spatial resolution and frame rate of the DMD. We expect, thus, the range of achievable $\Gamma$'s to 
be feasibly extended both to very high ($\Gamma \gg 1$) and very low ($\Gamma \ll 1$) values.}\\
\noindent {\it Conclusions.}
We have presented results, from numerical simulations and theoretical analysis, on the dewetting of thin liquid films on a switchable substrate, modelled with a space and time periodically varying contact angle in the thin-film equation.
Studying how the stability of the film depends on the underlying static pattern, we found that the rupture times grow linearly with the pattern wavelength, for short wavelengths, and quadratically in the long wavelength limit. 
In the time-dependent case, the rupture times are in general longer, indicating an induced greater film stability, and, while the quadratic growth is preserved at long wavelengths, a plateauing behaviour is observed as the wavelength decreases. 
A theoretical explanation has been provided for all these various regimes.
Furthermore, we showed that, at increasing the wettability wave speed, a transition occurs in the dewetting morphology from a multi-droplet to a metastable multi-rivulet state. 
A dimensionless parameter, $\Gamma$, controlling the transition has been identified in the ratio of the pattern speed and the typical film retraction speed over the substrate. 
The rivulets life-time itself grows with the pattern speed, displaying a logarithmic dependence that has been captured by means of phenomenological arguments.
On a broader perspective, our work suggests that switchable substrates offer a new avenue to control thin film dewetting, with obviously relevant implications, for instance, for open microfluidic devices, and paves the way to future studies in this direction, exploiting more complex and dedicated space-time dependencies.

\noindent {\it Acknowledgements}. SZ and JH acknowledge financial support from the Deutsche Forschungsgemeinschaft (DFG) within the priority program SPP2171 ``Dynamic Wetting of Flexible, Adaptive, and Switchable Substrates'', project HA-4382/11 and Project-ID 431791331—CRC1452 ``Catalysis at Liquid Interfaces''.
